FictionEro2 - Affective Impact of Fiction

### 2025-05-02 study1/analysis/2_emotions.qmd
updated analysis scripts and data

This model examines how Sex, Relevance, and Condition influence Valence ratings, treating relevance within each participant and items as random effects.

>>>>

This model examines how Sex, Relevance, and Condition influence Enticement ratings, treating relevance within each participant and items as random effects.

### 2025-05-02 study2/analysis/1_cleaning.qmd
updated analysis scripts and data

@@

```{r}
df |> 
  select(starts_with("Feedback"), -Feedback_Text, -Feedback_AllRealConfidence, -Feedback_AllFakeConfidence, -Feedback_DiffSubtle, -Feedback_Enjoyment) |>
  pivot_longer(everything(), names_to = "Question", values_to = "Answer") |>
  group_by(Question, Answer) |> 
  summarise(prop = n()/nrow(df), .groups = 'drop') |> 
  complete(Question, Answer, fill = list(prop = 0)) |> 
  filter(Answer == "True") |> 
  mutate(Question = str_remove(Question, "Feedback_"),
         Question = str_replace(Question, "SomeFacesAttractive", "Some faces were attractive"),
         Question = str_replace(Question, "NoFacesAttractive", "No faces were attractive"),
         Question = str_replace(Question, "AILessAttractive", "AI = Less attractive"),
         Question = str_replace(Question, "AIMoreAttractive", "AI = More attractive"),
         # Question = str_replace(Question, "DiffSubtle", "Subtle differences between AI images and photos"),
         Question = str_replace(Question, "DiffNone", "No difference between AI images and photos"),
         Question = str_replace(Question, "LabelsIncorrect", "Labels were incorrect"),
         Question = str_replace(Question, "AllReal", "All images were real"),
         Question = str_replace(Question, "AllFake", "All images were AI-generated"),
         Question = str_replace(Question, "DiffObvious", "Obvious differences between AI images and photos"),
         Question = str_replace(Question, "LabelsReversed", "Labels were reversed")) |>
  mutate(Question = fct_reorder(Question, desc(prop))) |> 
  ggplot(aes(x = Question, y = prop)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks(), labels=scales::percent) +
  labs(x="Feedback", y = "Participants", title = "Feedback") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 55, hjust = 1),
    axis.title.x = element_blank()
  )
```

>>>>

### Old vs New feedback

Due to a mistake in the study's implementation, some of the feedback was incorrect, e.g., "Some faces were attractive" and "AI-generated faces were less attractive" rather than "Some images were arousing" and "AI-generated images were less arousing".

table(df$version)

@@

```{r}
df |> 
  select(starts_with("Feedback"), -Feedback_Text, -Feedback_AllRealConfidence, -Feedback_AllFakeConfidence, -Feedback_DiffSubtle, -Feedback_Enjoyment) |>
  pivot_longer(everything(), names_to = "Question", values_to = "Answer") |>
  group_by(Question, Answer) |> 
  summarise(prop = n()/nrow(df), .groups = 'drop') |> 
  complete(Question, Answer, fill = list(prop = 0)) |> 
  filter(Answer == "True") |> 
  mutate(Question = str_remove(Question, "Feedback_"),
         Question = str_replace(Question, "SomeFacesAttractive", "Some faces were attractive"),
         Question = str_replace(Question, "NoFacesAttractive", "No faces were attractive"),
         Question = str_replace(Question, "AILessAttractive", "AI = Less attractive"),
         Question = str_replace(Question, "AIMoreAttractive", "AI = More attractive"),
         Question = str_replace(Question, "SomeImagesArousing", "Some images were arousing"),
         Question = str_replace(Question, "NoImagesArousing", "No images were arousing"),
         Question = str_replace(Question, "AILessArousing", "AI = Less arousing"),
         Question = str_replace(Question, "AIMoreArousing", "AI = More arousing"),
         # Question = str_replace(Question, "DiffSubtle", "Subtle differences between AI images and photos"),
         Question = str_replace(Question, "DiffNone", "No difference between AI images and photos"),
         Question = str_replace(Question, "LabelsIncorrect", "Labels were incorrect"),
         Question = str_replace(Question, "AllReal", "All images were real"),
         Question = str_replace(Question, "AllFake", "All images were AI-generated"),
         Question = str_replace(Question, "DiffObvious", "Obvious differences between AI images and photos"),
         Question = str_replace(Question, "LabelsReversed", "Labels were reversed")) |>
  mutate(Question = fct_reorder(Question, desc(prop))) |> 
  ggplot(aes(x = Question, y = prop)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks(), labels=scales::percent) +
  labs(x="Feedback", y = "Participants", title = "Feedback") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 55, hjust = 1),
    axis.title.x = element_blank()
  )
```

### 2025-05-02 study2/analysis/2_emotion.qmd
updated analysis scripts and data

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency, SexualActivity, BAIT_Worry, BAIT_Dangerous, 
           -Feedback_Text),
    by = "Participant"
    ) |> 
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexurualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0 ~ "True",
           Condition == "AI-Generated" & Realness < 0 ~ "True",
           .default = "False",
         )) |>
  mutate(across(starts_with("Feedback_")& !matches("Feedback_AllRealConfidence|Feedback_AllFakeConfidence|Feedback_Enjoyment"), function(x) {fct_relevel(x, "False", "True")}))

@@

bait <- df |> 
  select(starts_with("BAIT_"), -Bait_Duration) |> 
  rename_with(function(x) gsub("BAIT_\\d_", "", x)) |> 
  filter(!is.na(BAIT_TextRealistic ))

raw_scores <- data.frame(
  Participant = df[!is.na(df$BAIT_ImagesRealistic), ]$Participant,
  BAIT_Videos = (bait$BAIT_VideosRealistic + (1 - bait$BAIT_VideosIssues)) / 2,
  BAIT_Visual = (bait$BAIT_ImagesRealistic + (1 - bait$BAIT_ImagesIssues) + bait$BAIT_ImitatingReality + bait$BAIT_EnvironmentReal) / 4,
  BAIT_Text = (bait$BAIT_TextRealistic + (1 - bait$BAIT_TextIssues)) / 2
  )

df <- merge(df, raw_scores, by="Participant")
```

@@

@@

# marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender") |>
#   results_table()

@@

@@

moderator_results <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI", "PornFrequency", "SexualActivity")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Arousal", m))
}

@@

```{r}
feedback_results <- data.frame()
for(m in c("Feedback_NoFacesAttractive", "Feedback_SomeFacesAttractive", "Feedback_AIMoreAttractive", 
           "Feedback_AILessAttractive", "Feedback_DiffObvious", "Feedback_DiffSubtle", "Feedback_DiffNone" ,
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed", "Feedback_AllReal", "Feedback_AllFake",
           "Feedback_AllRealConfidence", "Feedback_AllFakeConfidence" ,  "Feedback_Enjoyment")){ 
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Arousal", m))
}
display(feedback_results)
```

@@

```{r}
  m <- glmmTMB::glmmTMB(Arousal ~ Sex / Condition + (1|Participant) + (1|Item),
                        data=filter(df, ConditionBelief == "True"), 
                        family=glmmTMB::ordbeta(),
                        control = glmmTMB::glmmTMBControl(parallel = 8))

@@

 moderator_believed <- data.frame()
  for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
             "GAAIS_Positive", "GAAIS_Negative",
             "Porn", "PornFrequency", "SexualActivity_num", 
             "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
             "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
             "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
    moderator_believed <- rbind(moderator_believed,
                               test_moderator(filter(df, ConditionBelief == "True"), "Arousal", m))
  }

@@

>>>>

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency, SexualActivity,- Feedback_Text),
    by = "Participant"
    ) |> 
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexurualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0 ~ "True",
           Condition == "AI-Generated" & Realness < 0 ~ "True",
           .default = "False",
         )) |>
  rename(AllRealConfidence = "Feedback_AllRealConfidence",
         AllFakeConfidence = "Feedback_AllFakeConfidence",
         Enjoyment = "Feedback_Enjoyment"
         )|>
  mutate(across(starts_with("Feedback_"), as.factor))

@@

df$BAIT_Videos <- (df$BAIT_VideosRealistic + (1 - df$BAIT_VideosIssues)) / 2

df$BAIT_Visual = (df$BAIT_ImagesRealistic + (1 - df$BAIT_ImagesIssues) + df$BAIT_ImitatingReality + df$BAIT_EnvironmentReal) / 4

df$BAIT_Text = (df$BAIT_TextRealistic + (1 - df$BAIT_TextIssues)) / 2

df$BAIT_Negative <- (df$BAIT_Worry + df$BAIT_Dangerous)/2

df$BAIT_Positive <- (df$BAIT_Exciting + df$BAIT_Benefit)/2
```

@@

The analysis also examined the average marginal effect of the variable "Condition" (comparing AI-Generated to Photograph) separately for males and females. For males, the arousal is on average significantly lower (althoug only slighly) for AI-Generated images compared to Photograph images. For females, the arousal is on average slightly lower for AI-Generated images compared to Photograph images; however, this effect is not significant.

@@

marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender") 
#   results_table()

@@

The analysis revealed a significant interaction effect only for **positive attitudes towards AI** among females. Specifically, females with higher positive attitudes towards AI rated AI-generated images as more arousing, relative to when they viewed photographs

No other moderators showed statistically significant effects on arousal across the interactions of gender, condition (Ai-generated vs Photograph), and the respective moderators.

We also examined whether **feedback** moderates the relationship between gender and condition on arousal. Findings suggest that feedback moderates this relationship only for men. Specifically, males who reported finding AI-generated images as less and more arousing had lower and higher, respectively, arousal ratings for those images.

<!-- All significant findings for feedback  -->

<!-- 1.For Males, giving feedback indicating that AI-generated images were less attractive, influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images as less arousing had lower arousal ratings for those images.

<!-- 2. For Males, giving feedback indicating that AI-generated images were more attractive significantly influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images more attractive had higher arousal ratings for these images.  -->

<!-- Note: The feedback that the difference between the conditions was subtle only had one value - i.e., False.  The feedback that some AI images were arousing and that no AI images were arousing also only had one value - i.e., False -->

@@

moderator_results <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI", "BAIT_Positive", "BAIT_Negative", "PornFrequency", "SexualActivity")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Arousal", m))
}

@@

```{r}
feedback_results <- data.frame()
for(m in c("Feedback_NoFacesAttractive", "Feedback_SomeFacesAttractive", "Feedback_AIMoreAttractive", "Feedback_AILessAttractive", "Feedback_DiffObvious", "Feedback_DiffNone" ,
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed", "Feedback_AllReal", "Feedback_AllFake", "AllRealConfidence", "AllFakeConfidence" ,  "Enjoyment",

           ## New feedback
           "Feedback_AIMoreArousing", "Feedback_AILessArousing"

           )){ 
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Arousal", m))
}
display(feedback_results)
```

@@

<!-- To DO -->

```{r}
  m <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition + (1|Participant) + (1|Stimulus),
                        data=filter(df, ConditionBelief == "True"), 
                        family=glmmTMB::ordbeta(),
                        control = glmmTMB::glmmTMBControl(parallel = 8))

@@

 moderator_believed <- data.frame()
  for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
             "BAIT_Positive", "BAIT_Negative",
             "PornFrequency", "SexualActivity", 
             "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
             "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
             "Feedback_AIMoreAttractive", "Feedback_AILessAttractive",
             "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
    moderator_believed <- rbind(moderator_believed,
                               test_moderator(filter(df, ConditionBelief == "True"), "Arousal", m))
  }

@@

## Enticement

Enticment ratings differ by whether the images were preceeded by 'photograph' or by 'ai-generated'.

enticement_desc <- plot_descriptive(df, "Enticing")
enticement_desc
```

This model examines how Gender and Condition influence enticement ratings, treating participant and items as random effects.

<!-- Key Findings: -->

There was a significant interaction between gender and image condition, whereby men were less enticed by AI-generated images, whereas women showed no significant difference in enticement based on image condition. Across both conditions, women rated the images as less enticing overall than men.

We also analyzed the average marginal effect of the variable “Condition” (comparing AI-Generated to Photograph) separately for males and females.

For males, enticement ratings are significantly lower on average for AI-Generated images compared to Photograph images. For females, enticement ratings are slightly lower on average for AI-Generated images compared to Photograph images, although this difference is not statistically significant.

**Note**. 'Other' was removed due to the very small percentage of individuals identifying as such.

```{r}
m_enticement <- glmmTMB::glmmTMB(Enticing ~ Gender / Condition + 
                                   (1 | Participant) + (1|Stimulus),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m_enticement)

marginaleffects::avg_slopes(m_enticement, variables="Condition", by="Gender", re.form=NA)
```

We examined whether **attitudes towards AI, porn frequency, and sexual activity** moderated the relationship between the predictors (sex, relevance, condition) and enticement ratings.

Looking at **feedback as a moderator**, the results showed that enticement ratings for images labeled AI-generated were moderated by the feedback reporting attitudes towards AI, but this was only significant for women. Specifically, women who reported that having more positive and negative attitudes towards AI, rated these images higher, and lower, in enticement compared to images labeled as photographs.

```{r}
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
           "BAIT_Positive", "BAIT_Negative",
           "PornFrequency", "SexualActivity")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Enticing", m))
}

display(filter(moderator_results, Outcome == "Enticing"))
```

```{r}
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_LessArousing",
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Enticing", m))
}
display(filter(feedback_results, Outcome == "Enticing"))
```

```{r}
m <- glmmTMB::glmmTMB(Enticing ~ Gender / (Condition * SexualActivity) + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

sex_act_plot<- estimate_relation(m, length=20) |> 
  ggplot(aes(x=SexualActivity_num, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.2) +
  geom_line(aes(color=Condition)) +
  facet_grid(~Sex)

sex_act_plot
```
